---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`askr` provides a minimal interface for calling Azure-hosted LLMs from R.

## Install and load

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("leadership-research/askr")
library(askr)
```

## Configure credentials

`askr` requires three environment variables. The recommended approach is to add
them to `~/.Renviron` so they are set automatically on every R session:

```
API_KEY=your-azure-api-token
ASKR_OPENAI_BASE=https://your-resource.cognitiveservices.azure.com
ASKR_OSS_BASE=https://your-resource.services.ai.azure.com
```

Then reload without restarting R:

```{r eval=FALSE}
readRenviron("~/.Renviron")
```

Or set them inline for the current session only:

```{r eval=FALSE}
Sys.setenv(
  API_KEY          = "your-azure-api-token",
  ASKR_OPENAI_BASE = "https://your-resource.cognitiveservices.azure.com",
  ASKR_OSS_BASE    = "https://your-resource.services.ai.azure.com"
)
```

## Run a first request

```{r eval=FALSE}
response <- ask(
  prompt = "Explain supervised vs unsupervised learning in one paragraph.",
  model = "gpt-4o",
  temperature = 0.4
)

cat(response)
```

## Custom system prompt

Override the default system prompt to give the model domain-specific context:

```{r eval=FALSE}
response <- ask(
  prompt        = "Identify the three main leadership themes in this text: [text]",
  model         = "mistral",
  system_prompt = "You are an expert in organisational psychology and leadership development."
)
```

## Extra API parameters

Pass additional parameters supported by the Azure API (e.g., `max_tokens`, `top_p`):

```{r eval=FALSE}
response <- ask(
  prompt       = "Summarise this document in bullet points.",
  model        = "gpt-4o",
  extra_params = list(max_tokens = 512L, top_p = 0.9)
)
```

## Programmatic usage

`ask_raw()` returns a structured `askr_response` object with status, model, and
response fields — useful for programmatic control or debugging:

```{r eval=FALSE}
res <- ask_raw(
  query         = "Extract 3 key risks from this text.",
  model         = "mistral",
  temperature   = 0.3,
  system_prompt = "You are a risk analyst.",
  extra_params  = list(max_tokens = 256L)
)

if (res$ok) {
  cat(res$response)
} else {
  message("HTTP status: ", res$status)
  message(res$response)
}
```

`print()` truncates output at 500 characters by default. Use `max_chars = Inf`
to see the full response:

```{r eval=FALSE}
print(res, max_chars = Inf)
```

## Discover available models

```{r eval=FALSE}
list_available_models()
```

## PII and sensitive data

**Do not send raw personally identifiable information as a prompt.**

The `query`/`prompt` argument is transmitted to an external Azure endpoint.
Before passing research data to `ask()` or `ask_raw()`, ensure it has been:

- De-identified (names, participant IDs, and contact details removed or pseudonymised)
- Reviewed under your organisation's data governance policy for suitability of external AI processing
- Checked against any IRB or ethics approval conditions that govern your study

If you work with interview transcripts or survey responses, pass a summary or
anonymised excerpt — not the raw text.

```{r eval=FALSE}
# Preferred: pass a pre-processed excerpt
ask(
  prompt = "Summarise the leadership themes in this anonymised excerpt: [excerpt]",
  model  = "mistral"
)

# Avoid: raw participant data in the prompt
# ask(prompt = raw_transcript_with_names, model = "mistral")  # don't do this
```

The `print()` method for `askr_response` objects truncates output at 500
characters by default to avoid inadvertently echoing large volumes of model
output to shared console sessions or log files. Use `print(res, max_chars = Inf)`
to see the full response.

For deeper implementation details, see the Architecture Overview vignette.
